---
title: "p8105_hw6_agg2965"
author: "Aakriti Shukla"
date: "2022-11-28"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(ggplot2)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

Problem 1
For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 0∗β̂ 1)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).

```{r}
weatherboot_df = 
  tibble(
    x = tmin,
    error = r̂ 2
    log(β̂ 0∗β1), #what is the error term here
    y = tmax
  )


boot_straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weatherboot_df))
  )

boot_straps

```

Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.

Let's load the data. 
```{r}
homicides_df <- read.csv(url("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"))%>%
  janitor::clean_names()
```

Next, I have created a city_state variable and a binary variable indicating whether the homicide is solved. I have omitted cities don’t report victim race and data entry mistakes. I have limtied the analysis those for whom victim_race is white or black and ensured that victim_age is numeric.
```{r}
homicides_clean=homicides_df%>%
  janitor::clean_names()%>%
    unite(city_state, c(city, state), sep = ", ")%>%
    filter(victim_race=="White" | victim_race=="Black")%>%
    subset(city_state!='Dallas, TX'& city_state!='Phoenix, AZ' & city_state!= 'Kansas City, MO' & city_state!= 'Tulsa, OK' & city_state!= 'Tulsa, AL')%>%
  mutate(
    solved = case_when(disposition=="Closed by arrest" ~ '1',
                           disposition=="Closed without arrest" |                              disposition=="Open/No arrest" ~ '0'),
    victim_age=as.numeric(victim_age),
    solved=as.numeric(solved),
    )

```
For the city of Baltimore, MD, we have used the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and 

Need to do this: obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_homicides=homicides_clean%>%
  filter(city_state %in% "Baltimore, MD")
```

```{r}
balt_fit = baltimore_homicides %>% glm(solved~victim_age + victim_sex + victim_race, data=., family=binomial())

balt_fit%>% 
  broom::glance()

summary(balt_fit)

balt_fit%>%
  broom::tidy()%>%
  mutate(OR=exp(estimate),
         OR_conf_low=exp(OR-1.96*(std.error)),
         OR_conf_high=exp(OR+1.96*(std.error)))%>%
    select(term,log_OR=estimate,OR, OR_conf_low, OR_conf_high)
```

Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
homicides_clean %>% 
  glm(solved~victim_age + victim_sex + victim_race, data = .) %>%
  broom::tidy() %>% 
  knitr::kable(digits = 3)

homicides_clean%>%
  group_by(city_state)%>%
    summarize(
      total=n())

homicides_fit =
  homicides_clean %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(.x=data, ~glm(solved~victim_age + victim_sex + victim_race, data = .x, family=binomial())),
    results = map(models, broom::tidy)) %>% 
  unnest(results)%>%
  filter(term == "victim_sexMale") %>%
    mutate(OR = exp(estimate),
           OR_conf_low = exp(estimate - 1.96*(std.error)),
         OR_conf_high = exp(estimate + 1.96*(std.error))) %>%
  select(city_state, OR, OR_conf_low, OR_conf_high) 

homicides_fit
```


Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.


```{r}
homicides_ordered=homicides_fit%>%
  mutate(
    city_state=fct_reorder(city_state,OR))
ggplot(homicides_ordered, aes(x=city_state, y=OR)) + 
    geom_errorbar(aes(ymin=OR_conf_low, ymax=OR_conf_high), width=.1) +
    geom_line() +
    geom_point()



```

```{r}
balt_fit%>%
  dplyr::select(starts_with("rmse"))%>%
  pivot_longer(
    everything(),
    names_to="model",
    values_to"rmse",
    names_prefix="rmse_")%>%
  mutate(model=fct_inorder(model))%>%
  ggplot(aes(x=model,y=rmse))+geom_violin()
```


Problem 3
In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

First, let's load the data.

```{r}
birthweight_df<- read.csv("data/birthweight.csv")%>%
  janitor::clean_names()
```

babysex: baby’s sex (male = 1, female = 2)
bhead: baby’s head circumference at birth (centimeters)
blength: baby’s length at birth (centimeteres)
bwt: baby’s birth weight (grams)
delwt: mother’s weight at delivery (pounds)
fincome: family monthly income (in hundreds, rounded)
frace: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
gaweeks: gestational age in weeks
malform: presence of malformations that could affect weight (0 = absent, 1 = present)
menarche: mother’s age at menarche (years)
mheigth: mother’s height (inches)
momage: mother’s age at delivery (years)
mrace: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
parity: number of live births prior to this pregnancy
pnumlbw: previous number of low birth weight babies
pnumgsa: number of prior small for gestational age babies
ppbmi: mother’s pre-pregnancy BMI
ppwt: mother’s pre-pregnancy weight (pounds)
smoken: average number of cigarettes smoked per day during pregnancy
wtgain: mother’s weight gain during pregnancy (pounds)
Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

