---
title: "p8105_hw6_agg2965"
author: "Aakriti Shukla"
date: "2022-11-28"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(ggplot2)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


#Problem 1

We used the the 2017 Central Park weather data for this problem. First, we'll load the data.


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```
To obtain a distribution for r̂^2$, we'll follow basically the same procedure we used for regression coefficients: draw bootstrap samples; the a model to each; extract the value I'm concerned with; and summarize. Here, we'll use `modelr::bootstrap` to draw the samples and `broom::glance` to produce `r.squared` values. 

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

In this example, the r̂^2 value is high, and the upper bound at 1 may be a cause for the generally skewed shape of the distribution. If we wanted to construct a confidence interval for r̂^2, we could take the 2.5% and 97.5% quantiles of the estimates across bootstrap samples. However, because the shape isn't symmetric, using the mean +/- 1.96 times the standard error probably wouldn't work well.

We can produce a distribution for $\log(\beta_0 * \beta1)$ using a similar approach, with a bit more wrangling before we make our plot.

```{r}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(
    names_from = term, 
    values_from = estimate) %>% 
  rename(beta0 = `(Intercept)`, beta1 = tmin) %>% 
  mutate(log_b0b1 = log(beta0 * beta1)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

As with $r^2$, this distribution is somewhat skewed and has some outliers. 



#Problem 2

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. We will perform logistic regression analysis on this dataset.

Let's load the data. 
```{r}
homicides_df <- read.csv(url("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"))%>%
  janitor::clean_names()
```

Next, I have created a city_state variable and a binary variable indicating whether the homicide is solved. I have omitted cities don’t report victim race and data entry mistakes. I have limtied the analysis those for whom victim_race is white or black and ensured that victim_age is numeric.
```{r}
homicides_clean=homicides_df%>%
  janitor::clean_names()%>%
    unite(city_state, c(city, state), sep = ", ")%>%
    filter(victim_race=="White" | victim_race=="Black")%>%
    subset(city_state!='Dallas, TX'& city_state!='Phoenix, AZ' & city_state!= 'Kansas City, MO' & city_state!= 'Tulsa, OK' & city_state!= 'Tulsa, AL')%>%
  mutate(
    solved = case_when(disposition=="Closed by arrest" ~ '1',
                           disposition=="Closed without arrest" |                              disposition=="Open/No arrest" ~ '0'),
    victim_age=as.numeric(victim_age),
    solved=as.numeric(solved),
    )

```
For the city of Baltimore, MD, we have used the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. We have obtained an estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_homicides=homicides_clean%>%
  filter(city_state %in% "Baltimore, MD")
```

```{r}
balt_fit = baltimore_homicides %>% glm(solved~victim_age + victim_sex + victim_race, data=., family=binomial())

balt_fit%>% 
  broom::glance()

summary(balt_fit)

balt_fit%>%
  broom::tidy()%>%
  mutate(OR=exp(estimate),
         OR_conf_low=exp(OR-1.96*(std.error)),
         OR_conf_high=exp(OR+1.96*(std.error)))%>%
    select(term,log_OR=estimate,OR, OR_conf_low, OR_conf_high)
```

We then ran a glm for each of the cities in the dataset and extracted the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.

```{r}
homicides_clean %>% 
  glm(solved~victim_age + victim_sex + victim_race, data = .) %>%
  broom::tidy() %>% 
  knitr::kable(digits = 3)

homicides_clean%>%
  group_by(city_state)%>%
    summarize(
      total=n())

homicides_fit =
  homicides_clean %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(.x=data, ~glm(solved~victim_age + victim_sex + victim_race, data = .x, family=binomial())),
    results = map(models, broom::tidy)) %>% 
unnest(results)%>%
  filter(term == "victim_sexMale") %>%
    mutate(OR = exp(estimate),
           OR_conf_low = exp(estimate - 1.96*(std.error)),
         OR_conf_high = exp(estimate + 1.96*(std.error))) %>%
  select(city_state, OR, OR_conf_low, OR_conf_high) 

homicides_fit
```


Below, we have created a plot that shows the estimated ORs and CIs for each city. The odds of solving homicides of male victims was lower than the odds of solving homicides of female victims in all studied cities except for in Fresno CA, Stockton, CA, and Albuquerque, NM.


```{r}
homicides_ordered=homicides_fit%>%
  mutate(
    city_state=fct_reorder(city_state,OR))
ggplot(homicides_ordered, aes(x=city_state, y=OR)) + 
    geom_errorbar(aes(ymin=OR_conf_low, ymax=OR_conf_high), width=.1) +
    geom_line() +
    geom_point()+
    labs(title = "Odds of solving homicides of male vs. female victims",
           x = "City and State",
       y="Adjusted OR") +
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



#Problem 3

In this problem, we analyzed data gathered to understand the effects of several variables on a child’s birthweight.

First, let's load the data.

```{r}
birthweight_df<- read.csv("data/birthweight.csv")%>%
  janitor::clean_names()%>%
  mutate(
    id=1:4342
  )
```

Below, I have proposed a regression model (Model 1) for birthweight. This model included  head circumference, length, and gestational age in weeks as predictors of birthweight. I first created this model and made a plot of model residuals against fitted values using add_predictions and add_residuals. The scatterplot demonstrates clustering from 2000g to 4000g, with a few outlying points in the very low birthweight range, suggesting that this model may not be as valid for very low birthweight babies. Of note, there are 3 negative birthweight values, which appear to be errors in documentation.


```{r}
bwt_fit1 = birthweight_df %>% lm (bwt~bhead+blength+gaweeks, data=.)

bwt_fit1%>% 
  broom::glance()

summary(bwt_fit1)

birthweight_df %>% 
  modelr::add_residuals(bwt_fit1) %>% 
  modelr::add_predictions(bwt_fit1) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()

```

I have now compared this model to two others. Model 2 uses length at birth and gestational age as predictors, without including interaction effects. The scatterplot demonstrates clustering from 2000g to 4000g.  This plot also has a a few outlying points in the very low birthweight range, suggesting that this model may not be as valid for very low birthweight babies. Of note, there is 1 negative birthweight value, which appears to be an error in documentation.

```{r}
bwt_fit2 = birthweight_df %>% lm (bwt~blength+gaweeks, data=.)

bwt_fit2%>% 
  broom::glance()

summary(bwt_fit2)

birthweight_df  %>% 
  modelr::add_residuals(bwt_fit2) %>% 
  modelr::add_predictions(bwt_fit2) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()
```


The other model (Model 3) uses head circumference, length, sex, and all interactions (including the three-way interaction) between these variables. The associated scatterplot shows findings as the other models. 

```{r}
bwt_fit3 = birthweight_df %>% lm (bwt~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data=.)

bwt_fit3%>% 
  broom::glance()

summary(bwt_fit3)

birthweight_df  %>% 
  modelr::add_residuals(bwt_fit3) %>% 
  modelr::add_predictions(bwt_fit3) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()
```


Next, we compare the three models using cross-validated prediction error as per below. We have created resulting violin plots, which show that the root mean standard error is highest with the second model, which includes only length at birth and gestational age in weeks. This suggests that this model is inferior to the others. This is unsurprising as models 1 and 3 contained more prediction variables as compared to model 2. 
```{r}
cv_df =
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    bwt_fit1  = map(train, ~lm(bwt~ bhead+blength+gaweeks, data=.x)),
    bwt_fit2  = map(train, ~lm(bwt~blength+gaweeks, data=.x)),
    bwt_fit3  = map(train, ~lm(bwt~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data=.x))) %>% 
  mutate(
    rmse_fit1 = map2_dbl(bwt_fit1, test, ~rmse(model = .x, data = .y)),
    rmse_fit2 = map2_dbl(bwt_fit2, test, ~rmse(model = .x, data = .y)),
    rmse_fit3 = map2_dbl(bwt_fit3, test, ~rmse(model = .x, data = .y)))

cv_df%>% 
    select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin()

```

