---
title: "p8105_hw6_agg2965"
author: "Aakriti Shukla"
date: "2022-11-28"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)
library(ggplot2)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

This homework includes figures; the readability of your embedded plots (e.g. font sizes, axis labels, titles) will be assessed in Problems 1+.

Problem 1
For this problem, we’ll use the 2017 Central Park weather data that we’ve seen elsewhere. The code chunk below (adapted from the course website) will download these data.


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```


The boostrap is helpful when you’d like to perform inference for a parameter / value / summary that doesn’t have an easy-to-write-down distribution in the usual repeated sampling framework. We’ll focus on a simple linear regression with tmax as the response and tmin as the predictor, and are interested in the distribution of two quantities estimated from these data:

r̂ 2
log(β̂ 0∗β̂ 1)
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. Plot the distribution of your estimates, and describe these in words. Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r̂ 2 and log(β̂ 0∗β̂ 1). Note: broom::glance() is helpful for extracting r̂ 2 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β̂ 0∗β̂ 1).

```{r, eval=FALSE}
weatherboot_df = 
  tibble(
    x = tmin,
    error = r̂ 2
    log(β̂ 0∗β1), #what is the error term here
    y = tmax
  )


boot_straps = 
  data_frame(
    strap_number = 1:5000,
    strap_sample = rerun(5000, boot_sample(weatherboot_df))
  )

boot_straps

```

Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. We will perform logistic regression analysis on this dataset.

Let's load the data. 
```{r}
homicides_df <- read.csv(url("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"))%>%
  janitor::clean_names()
```

Next, I have created a city_state variable and a binary variable indicating whether the homicide is solved. I have omitted cities don’t report victim race and data entry mistakes. I have limtied the analysis those for whom victim_race is white or black and ensured that victim_age is numeric.
```{r}
homicides_clean=homicides_df%>%
  janitor::clean_names()%>%
    unite(city_state, c(city, state), sep = ", ")%>%
    filter(victim_race=="White" | victim_race=="Black")%>%
    subset(city_state!='Dallas, TX'& city_state!='Phoenix, AZ' & city_state!= 'Kansas City, MO' & city_state!= 'Tulsa, OK' & city_state!= 'Tulsa, AL')%>%
  mutate(
    solved = case_when(disposition=="Closed by arrest" ~ '1',
                           disposition=="Closed without arrest" |                              disposition=="Open/No arrest" ~ '0'),
    victim_age=as.numeric(victim_age),
    solved=as.numeric(solved),
    )

```
For the city of Baltimore, MD, we have used the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. We have obtained an estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_homicides=homicides_clean%>%
  filter(city_state %in% "Baltimore, MD")
```

```{r}
balt_fit = baltimore_homicides %>% glm(solved~victim_age + victim_sex + victim_race, data=., family=binomial())

balt_fit%>% 
  broom::glance()

summary(balt_fit)

balt_fit%>%
  broom::tidy()%>%
  mutate(OR=exp(estimate),
         OR_conf_low=exp(OR-1.96*(std.error)),
         OR_conf_high=exp(OR+1.96*(std.error)))%>%
    select(term,log_OR=estimate,OR, OR_conf_low, OR_conf_high)
```

We then ran a glm for each of the cities in the dataset and extracted the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.

```{r}
homicides_clean %>% 
  glm(solved~victim_age + victim_sex + victim_race, data = .) %>%
  broom::tidy() %>% 
  knitr::kable(digits = 3)

homicides_clean%>%
  group_by(city_state)%>%
    summarize(
      total=n())

homicides_fit =
  homicides_clean %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = map(.x=data, ~glm(solved~victim_age + victim_sex + victim_race, data = .x, family=binomial())),
    results = map(models, broom::tidy)) %>% 
unnest(results)%>%
  filter(term == "victim_sexMale") %>%
    mutate(OR = exp(estimate),
           OR_conf_low = exp(estimate - 1.96*(std.error)),
         OR_conf_high = exp(estimate + 1.96*(std.error))) %>%
  select(city_state, OR, OR_conf_low, OR_conf_high) 

homicides_fit
```


Below, we have created a plot that shows the estimated ORs and CIs for each city. The odds of solving homicides of male victims was lower than the odds of solving homicides of female victims in all studied cities except for in Fresno CA, Stockton, CA, and Albuquerque, NM.


```{r}
homicides_ordered=homicides_fit%>%
  mutate(
    city_state=fct_reorder(city_state,OR))
ggplot(homicides_ordered, aes(x=city_state, y=OR)) + 
    geom_errorbar(aes(ymin=OR_conf_low, ymax=OR_conf_high), width=.1) +
    geom_line() +
    geom_point()+
    labs(title = "Odds of solving homicides of male vs. female victims",
           x = "City and State",
       y="Adjusted OR") +
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


Problem 3
In this problem, we analyzed data gathered to understand the effects of several variables on a child’s birthweight.

First, let's load the data.

```{r}
birthweight_df<- read.csv("data/birthweight.csv")%>%
  janitor::clean_names()%>%
  mutate(
    id=1:4342
  )
```

Below, I have proposed a regression model for birthweight. This model included  head circumference, length, and gestational age in weeks as predictors of birthweight. I first created this model and made a plot of model residuals against fitted values using add_predictions and add_residuals. The plot demonstrates clustering from 2000g to 4000g, with a few outlying points in the very low birthweight range, suggesting that this model may not be as valid for very low birthweight babies. Of note, there are 3 negative birthweight values, which appear to be errors in documentation.


```{r}
bwt_fit1 = birthweight_df %>% lm (bwt~bhead+blength+gaweeks, data=.)

bwt_fit1%>% 
  broom::glance()

summary(bwt_fit1)

birthweight_df %>% 
  modelr::add_residuals(bwt_fit1) %>% 
  modelr::add_predictions(bwt_fit1) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point()

```

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)

```{r}
bwt_fit2 = birthweight_df %>% lm (bwt~blength+gaweeks, data=.)

bwt_fit2%>% 
  broom::glance()

summary(bwt_fit2)
```


One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

```{r}
bwt_fit3 = birthweight_df %>% lm (bwt~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data=.)

bwt_fit3%>% 
  broom::glance()

summary(bwt_fit3)
```

```{r}
cv_df =
  crossv_mc(birthweight_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    bwt_fit1  = map(train, ~lm(bwt~ bhead+blength+gaweeks, data=.x)),
    bwt_fit2  = map(train, ~lm(bwt~blength+gaweeks, data=.x)),
    bwt_fit3  = map(train, ~lm(bwt~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data=.x))) %>% 
  mutate(
    rmse_fit1 = map2_dbl(bwt_fit1, test, ~rmse(model = .x, data = .y)),
    rmse_fit2 = map2_dbl(bwt_fit2, test, ~rmse(model = .x, data = .y)),
    rmse_fit3 = map2_dbl(bwt_fit3, test, ~rmse(model = .x, data = .y)))

cv_df%>% 
    select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
    geom_violin()

```

